---
title: "Timestep Data Analyses"
author: "Paige Amos"
date: "12/10/2021"
output: pdf_document
---

#### Load required libraries
```{r, message = FALSE, warning = FALSE}
library(here)
library(reshape2)
library(FSA) 
library(fGarch)
library(LambertW)
library(patchwork)
library(DHARMa)
library(mgcv)
library(fitdistrplus)
library(goft)
library(gamlss)
library(rstatix)
library(performance) 
library(scales)
library(viridis)
library(ggpubr)
library(tidyverse)
```

# Manipulate Raw Tube Growth Data for Plotting and Stats

## Load raw data
```{r, message = FALSE, warning = FALSE}
tube_growth = read_csv(here("./data/tube-growth.csv"))
knitr::kable(tube_growth, caption = "Tube growth rates of all marked worms in various treatments.")
```

## Wraggle data
```{r, message = FALSE, warning = FALSE}
# Clean tube_growth data for plotting and statistical tests
clean_tube_growth <- tube_growth %>%
  mutate(zone = as.factor(zone)) %>%
  unite("replicate", tank:zone, remove = TRUE) %>%
  dplyr::select(replicate, worm_id, amt_cut_mm, treatment, growth_1, growth_2, growth_3, growth_4, growth_5) %>%
  pivot_longer(!c(replicate, worm_id, amt_cut_mm, treatment),
               names_to = "timestep",
               values_to = "growth") %>%
  mutate(treatment = as.factor(treatment),
         worm_id = as.factor(worm_id),
         timestep = as.factor(timestep),
         growth = as.double(growth))

```

# Statistical Analyses

## Checking Assumptions of a One-way Repeated Measures ANOVA
```{r, message = FALSE, warning = FALSE}
# visualize data
ggplot(clean_tube_growth, aes(x= timestep, y = growth)) +
  geom_boxplot()

# check for normality of the growth at each timestep using Shapiro Wilks test
clean_tube_growth %>%
  group_by(timestep) %>%
  shapiro_test(growth)

# investigate distribution of the growth at each timestep using qqplots since shapiro-wilk test found growth at all timesteps to be not normal

# timestep 1
growth_1 <- clean_tube_growth %>%
  dplyr::filter(timestep == "growth_1")

qqnorm(growth_1$growth)
qqline(growth_1$growth)

# timestep 2
growth_2 <- clean_tube_growth %>%
  dplyr::filter(timestep == "growth_2")

qqnorm(growth_2$growth)
qqline(growth_2$growth)

# timestep 3
growth_3 <- clean_tube_growth %>%
  dplyr::filter(timestep == "growth_3")

qqnorm(growth_3$growth)
qqline(growth_3$growth)

# timestep 4
growth_4 <- clean_tube_growth %>%
  dplyr::filter(timestep == "growth_4")

qqnorm(growth_4$growth)
qqline(growth_4$growth)

# timestep 5
growth_5 <- clean_tube_growth %>%
  dplyr::filter(timestep == "growth_5")

qqnorm(growth_5$growth)
qqline(growth_5$growth)

# Since sample size is N=90 overall and n= 30 for each treatment, sample size is large enough to ignore the assumption of normality for the ANOVA

# Checking for extreme outliers
clean_tube_growth %>%
  group_by(timestep) %>%
  identify_outliers(growth)

# Since two extreme outliers were found (B_2_6, C_2_1) remove extreme outliers so that results of ANOVA with and without extreme outliers can be compared to determine if extreme outliers significantly affect results
no_exouts <- clean_tube_growth %>%
  dplyr::filter(growth < 7.90)

#check that extreme outliers were successfully removed
no_exouts %>%
  group_by(timestep) %>%
  identify_outliers(growth)

#assumption of sphericity automatically checked by ANOVA test

```

## One-way Repeated Measures Analysis of Variance (ANOVA)
```{r, message = FALSE, warning = FALSE}
# run one-way repeated measures ANOVA and obtain ANOVA table
tube.aov <-anova_test(data = clean_tube_growth, dv = growth,
                       wid = worm_id,
                       within = timestep)
get_anova_table(tube.aov)

# run one-way repeated measures ANOVA for data with extreme outliers removed

noexouts.aov <- anova_test(data = no_exouts, dv = growth,
                       wid = worm_id,
                       within = timestep)
get_anova_table(noexouts.aov)

# results of ANOVA do not change significantly (both are statistically significant) when extreme outliers are removed so results of ANOVA including extreme outliers will be reported
```

## Pairwise Comparison
```{r, message = FALSE, warning = FALSE}
# performing pairwise comparison between timesteps
pwc <- clean_tube_growth %>%
  pairwise_t_test(growth ~ timestep, paired = TRUE,
                  p.adjust.method = "bonferroni")

pwc <- pwc %>% add_xy_position(x = "timestep")

pwc

```

## Finding the Mean and Standard Error of Tube Growth in each Timestep
```{r, message = FALSE, warning = FALSE}
# finding the mean and standard error of tube growth in timestep 1
summary(tube_growth$growth_1)
se(tube_growth$growth_1)

# finding the mean and standard error of tube growth in timestep 2
summary(tube_growth$growth_2)
se(tube_growth$growth_2)

# finding the mean and standard error of tube growth in timestep 3
summary(tube_growth$growth_3)
se(tube_growth$growth_3)

# finding the mean and standard error of tube growth in timestep 4
summary(tube_growth$growth_4)
se(tube_growth$growth_4)

# finding the mean and standard error of tube growth in timestep 5
summary(tube_growth$growth_5)
se(tube_growth$growth_5)
```

## Finding the Mean and Standard Error of Tube Growth for Each Treatment Within all Timesteps
```{r, message = FALSE, warning = FALSE}
# Making a dataset containing only control growth values

control <- tube_growth %>%
  dplyr::filter(treatment == "control")

# making a dataset containing only low salinity treatment growth values

salinity <- tube_growth %>%
  dplyr::filter(treatment == "low_salinity")

# making a dataset containing only high temperature treatment growth values

temperature <- tube_growth %>%
  dplyr::filter(treatment == "high_temp")

# finding the mean and standard error of tube growth of all treatments for timestep 1

summary(control$growth_1)
se(control$growth_1)

summary(salinity$growth_1)
se(salinity$growth_1)

summary(temperature$growth_1)
se(temperature$growth_1)

# finding the mean and standard error of tube growth of all treatments for timestep 2

summary(control$growth_2)
se(control$growth_2)

summary(salinity$growth_2)
se(salinity$growth_2)

summary(temperature$growth_2)
se(temperature$growth_2)

# finding the mean and standard error of tube growth of all treatments for timestep 3

summary(control$growth_3)
se(control$growth_3)

summary(salinity$growth_3)
se(salinity$growth_3)

summary(temperature$growth_3)
se(temperature$growth_3)

# finding the mean and standard error of tube growth of all treatments for timestep 4

summary(control$growth_4)
se(control$growth_4)

summary(salinity$growth_4)
se(salinity$growth_4)

summary(temperature$growth_4)
se(temperature$growth_4)

# finding the mean and standard error of tube growth of all treatments for timestep 5

summary(control$growth_5)
se(control$growth_5)

summary(salinity$growth_5)
se(salinity$growth_5)

summary(temperature$growth_5)
se(temperature$growth_5)
```

# Plotting the Timestep Data
```{r, message = FALSE, warning = FALSE}
# choosing colours for one.way plot

one.way.cols = c("growth_1" = "#46337EFF", "growth_2" = "#365C8DFF", "growth_3" = "#277F8EFF", "growth_4" = "#1FA187FF", "growth_5" = "#9FDA3AFF")

# Making a plot of the mean growth of all treatments over all timesteps

one.way = 
  ggplot(clean_tube_growth, aes(x = timestep, y = growth)) +
  geom_boxplot(aes(fill = timestep)) +
  theme_classic() +
  theme(legend.position = "none") +
  xlab("Timestep") +
  ylab("Tube Growth (mm)") +
  scale_x_discrete(labels=c("growth_1" = "1", "growth_2" = "2","growth_3" = "3", "growth_4" = "4", "growth_5" = "5")) +
  scale_color_manual(values = one.way.cols) +
  scale_fill_manual(values = one.way.cols)

one.way

# saving one.way plot

ggsave(plot = one.way, filename = "timestep-growth.jpg")

#adding significance bars from pairwise comparisons to one.way plot (significant bars labelled ns were removed manually outside of r for the report)

p.one.way <- one.way  +
  stat_pvalue_manual(pwc) +
  labs(subtitle = get_test_label(tube.aov, detailed = TRUE),
    caption = get_pwc_label(pwc))

p.one.way

# saving p.one.way plot

ggsave(plot = p.one.way, filename = "p-value-timestep-growth.jpg", width = 6, height = 7, dpi = 200, units = "in")

# Making a plot of the mean tube growth per treatment over all timesteps

treat.growth =
  ggplot(clean_tube_growth, aes(x = timestep, y = growth, fill = treatment)) +
    geom_boxplot() +
    scale_fill_manual(name = "Treatment", labels = c("Control", "Temperature", "Salinity"), values = c("#46337EFF", "#277F8EFF", "#9FDA3AFF")) +
    theme_classic() +
    xlab("Timestep") +
    ylab("Tube Growth (mm)") +
    scale_x_discrete(labels=c("growth_1" = "1", "growth_2" = "2","growth_3" = "3", "growth_4" = "4", "growth_5" = "5")) 
    
treat.growth

# saving treat.growth plot

ggsave(plot = treat.growth, filename = "treatment-timestep-growth.jpg")

```


